Inducing Domain Specific Languages for Bayesian Program Learning

### 2018-02-05 paper.tex
attempt to redo the lower bound section

We frame this problem as maximum a posteriori (MAP) inference in the generative model diagrammed in Figure~\ref{graphicalModel}. We wish to maximize the posterior probability of $\mathcal{D}$:
\begin{equation*}
  \probability[\mathcal{D}|X]\propto \probability[\mathcal{D}] \int \mathrm{d}\theta P(\theta|\mathcal{D})\prod_{x\in X} \sum_p \probability[x|p]\probability[p|\mathcal{D},\theta]
\end{equation*}
In general this marginalization over $\theta$ is intractable, so we make an AIC-style approximation\footnote{Sec.~\ref{grammarInductionSection} explains that $\mathcal{D}$ is a context-sensitive grammar.
Conventional natural-language processing (NLP) approaches to using variational inference to lower bound the marginal over $\theta$ do not apply in our setting.}, $A\approx \log\probability[\mathcal{D}|X] $:
\begin{align}
  A =   \log \probability[\mathcal{D}] + \argmax_{\theta}& \sum_{x\in X}\log \sum_p\probability[x|p]\probability[p|\mathcal{D},\theta]\nonumber\\
&+  \log P(\theta|\mathcal{D}) - ||\theta||_0 \label{AIC}
  \end{align}

If we had a $(\mathcal{D},\theta)$ maximizing Eq.~\ref{AIC}, then we could recover the most likely program for task $x$ by maximizing $\probability[x|p] \probability[p|\mathcal{D},\theta]$.
Through this lens we now take as our goal to maximize Eq.~\ref{AIC}.
But even \emph{evaluating} Eq.~\ref{AIC} is intractable because it involves summing over the infinite set of all possible programs. In general, programs are hard-won: finding even a single program that explains a given observation presents a daunting combinatorial search problem.
With this fact in mind, we will instead maximize the following tractable lower bound on Eq.~\ref{AIC},
which we call $J$:
\begin{equation}
J = \log \probability[\mathcal{D},\theta] + \sum_{x\in X}\log \sum_{p\in \mathcal{F}_x} \probability[x|p]\probability[p|\mathcal{D},\theta]\label{lowerBound}
\end{equation}
This lower bound depends on sets of programs, $\left\{\mathcal{F}_x \right\}_{x\in X}$:

\noindent\textbf{Definition.} A \emph{frontier of task $x$}, written $\mathcal{F}_x$,
is a finite set of programs where $\probability[x|p] > 0$ for all $p\in \mathcal{F}_x$.

>>>>

We frame this problem as maximum a posteriori (MAP) inference in the generative model diagrammed in Figure~\ref{graphicalModel}(a), finding the $\mathcal{D}^*,\theta^*$ solving:
\begin{align}\label{intractableObjectives}
  \nonumber\probability[X|\mathcal{D},\theta] &= \prod_{x\in X} \sum_p \probability[x|p]\probability[p|\mathcal{D},\theta]\\
  \nonumber\probability[\mathcal{D}|X]&\propto \probability[\mathcal{D}] \int \mathrm{d}\theta P(\theta|\mathcal{D})\probability[X|\mathcal{D},\theta]\\
  \mathcal{D}^* &= \argmax_{\mathcal{D}}\probability[\mathcal{D}|X]\\
\nonumber  \theta^*& =\argmax_\theta P(\theta |\mathcal{D}^*)\probability[X|\mathcal{D}^*,\theta]
\end{align}
The above equations summarize the problem from the point of view of an ideal Bayesian learner.
However, Eq.~\ref{intractableObjectives}
is wildly intractable because evaluating $\probability[X|\mathcal{D},\theta]$ involves
summing over the  infinite set of all possible programs.
In practice we will only ever be able to sum over a finite set of programs.
So, for each task, we define a finite set of programs, called a \emph{frontier}, and only marginalize over the frontiers:
\\\noindent\textbf{Definition.} A \emph{frontier of task $x$}, written $\mathcal{F}_x$,
is a finite set of programs where $\probability[x|p] > 0$ for all $p\in \mathcal{F}_x$.

With the frontiers in hand, we can define the following intuitive lower bound on the likelihood of the tasks, which we call $\mathcal{L}$:
\begin{align}
  \probability[X|\mathcal{D},\theta]\geq J\triangleq\prod_{x\in X} \sum_{p\in \mathcal{F}_x} \probability[x|p]\probability[p|\mathcal{D},\theta]
\end{align}
along with a corresponding lower bound on the marginal likelihood of a DSL, which we call $$
\begin{align}
  \probability[\mathcal{D}|X]\geq 
  \end{align}

If we had a $(\mathcal{D},\theta)$ solving Eq.~\ref{intractableObjectives}, then we could recover the most likely program for task $x$ by maximizing $\probability[x|p] \probability[p|\mathcal{D},\theta]$.
Through this lens we now take as our goal to solve Eq.~\ref{intractableObjectives}.
But even \emph{evaluating} Eq.~\ref{intractableObjectives} is intractable because it involves summing over the infinite set of all possible programs, as an ideal Bayesian learner would.
In practice, we must instead marginalize over
some finite set of programs.

In general, programs are hard-won: finding even a single program that explains a given observation presents a daunting combinatorial search problem.

Now in theory we would like to sum over the entire infinite space
of all programs -- but this is of course impossible.

%% In general this marginalization over $\theta$ is intractable, so we make an AIC-style approximation\footnote{Sec.~\ref{grammarInductionSection} explains that $\mathcal{D}$ is a context-sensitive grammar.
%% Conventional natural-language processing (NLP) approaches to using variational inference to lower bound the marginal over $\theta$ do not apply in our setting.}, $A\approx \log\probability[\mathcal{D}|X] $:
%% \begin{align}
%%   A =   \log \probability[\mathcal{D}] + \argmax_{\theta}& \sum_{x\in X}\log \sum_p\probability[x|p]\probability[p|\mathcal{D},\theta]\nonumber\\
%% &+  \log P(\theta|\mathcal{D}) - ||\theta||_0 \label{AIC}
%%   \end{align}

With this fact in mind, we will instead maximize the following tractable lower bound on Eq.~\ref{AIC},
which we call $J$:
\begin{equation}
J = \log \probability[\mathcal{D},\theta] + \sum_{x\in X}\log \sum_{p\in \mathcal{F}_x} \probability[x|p]\probability[p|\mathcal{D},\theta]\label{lowerBound}
\end{equation}
This lower bound depends on sets of programs, $\left\{\mathcal{F}_x \right\}_{x\in X}$:
